---
title: "BRCA1 and Relative Solvent Accessibility"
author: "Melissa Cline and Diego Quezada-Munoz"
date: "May 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data Exploration

We read in the data from the CSV File that was produce by the previous parts in this project.
```{r}
data = read.csv(file.choose()) # Select brca1.csv
```


Here are the first 15 instances of the "brca1.csv" data set.
```{r}
head(data, 15)
```

The number of variants in this data set is 3338.
```{r}
nrow(data)
```

## Data Exploration

In the following long output we see all the classifications provided by ClinVar.
```{r}
table(data$clinvar)
```

In this code we exclude variants that have contradicting and unknown classifications.
```{r}
d = data[data$clinvar != 'variant_of_unknown_significance' & data$clinvar != 'Uncertain_significance' & data$clinvar != 'variant_of_unknown_significance,Uncertain_significance' & data$clinvar != 'Uncertain_significance,not_provided' & data$clinvar != 'Pathogenic,Benign,Uncertain_significance,Likely_benign' & data$clinvar != 'not_provided' & data$clinvar != 'no_known_pathogenicity' & data$clinvar != 'Likely_pathogenic,Likely_benign,Uncertain_significance' & data$clinvar != 'Likely_pathogenic,Benign,Uncertain_significance,not_provided,probably_not_pathogenic,Likely_benign',]
```


We then have 1,572 variants remaining. 
```{r}
nrow(d)
```

we are able to pickout the variants with a benign classification. There is a total of 1008 benign variants.
```{r}
b = grepl('enign', d$clinvar)
benign = d[b,]
nrow(benign)
rownames(benign) <- seq(length=nrow(benign))
```

There is a total of 564 pathogenic variants. 
```{r}
p = !b
pathogenic = d[p,]
nrow(pathogenic)
rownames(pathogenic) <- seq(length=nrow(pathogenic))
```


```{r}
benign$pPrior2 = as.numeric(levels(benign$pPrior2))[benign$pPrior2]
str(benign)
```


```{r}
pathogenic$pPrior2 = as.numeric(levels(pathogenic$pPrior2))[pathogenic$pPrior2]
str(pathogenic)
```


```{r}
benign = benign[!grepl("X", benign$amino_acid_change),]
nrow(benign)
```


```{r}
pathogenic = pathogenic[!grepl("X",pathogenic$amino_acid_change),]
nrow(pathogenic)
```


```{r}
bPath = benign$pPrior2 > .2
sum(bPath, na.rm=TRUE)
```

```{r}
bBen = benign$pPrior2 < .05
sum(bBen, na.rm=TRUE)
```


```{r}
pPath = pathogenic$pPrior2 > .2
sum(pPath, na.rm=TRUE)

```


```{r}
pBen = pathogenic$pPrior2 < .05
sum(pBen, na.rm=TRUE)

```


```{r}
par(mfrow = c(2, 2))
boxplot(benign[bBen,]$rsa, main="ClinVar: benign | UCSC: benign", ylim=c(0,1), ylab = "RSA Score")
boxplot(benign[bPath,]$rsa, main="ClinVar: benign | UCSC: pathogenic", ylim=c(0,1), ylab = "RSA Score")
boxplot(pathogenic[pBen,]$rsa, main="ClinVar: pathogenic | UCSC: benign", ylim=c(0,1), ylab = "RSA Score")
boxplot(pathogenic[pPath,]$rsa, main="ClinVar: pathogenic | UCSC: pathogenic", ylim=c(0,1), ylab = "RSA Score")
```

```{r}
medians = matrix(c(
quantile(benign[bBen,]$rsa, c(.5), na.rm = TRUE),
quantile(benign[bPath,]$rsa, c(.5), na.rm = TRUE),
quantile(pathogenic[pBen,]$rsa, c(.5), na.rm = TRUE),
quantile(pathogenic[pPath,]$rsa, c(.5), na.rm = TRUE)), 2, 2)
medians
```


```{r}
par(mfrow = c(1, 2))
boxplot(benign$rsa, main="ClinVar: benign", ylim=c(0,1), ylab = "RSA Score")
boxplot(pathogenic$rsa, main="ClinVar: pathogenic", ylim=c(0,1), ylab = "RSA Score")
```


```{r}
summary(benign$rsa)
```


```{r}
summary(pathogenic$rsa)
```
